---
title: "Design and sequential optimization"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Design and sequential optimization}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

The development version of `magp` can connect three stages of an experiment:

1. generate an initial quantitative-sequence design;
2. evaluate the response at those initial runs;
3. use expected improvement to select later runs one at a time.

The objective function in this article is deterministic and inexpensive. In a
real study, the same function interface may call a simulator, run a laboratory
procedure through an external system, or return a measured response entered
after an experiment is complete.

## Build an initial design

```{r initial-design}
library(magp)

design <- magp_initial_design(
  n = 8,
  q = 3,
  sequence_maxit = 200,
  quantity_maxit = 200,
  alignment_maxit = 200,
  seed = 4
)

design$design
design$criteria
```

The first three columns contain the quantitative levels. The last three
columns contain sequence positions. Each quantitative column remains a Latin
hypercube, and every sequence row remains a valid permutation.

## Define an objective

The integrated interface passes one named argument for every input column.
This example rewards quantitative levels near a target and a preferred
sequence.

```{r objective}
objective <- function(quantity_1, quantity_2, quantity_3,
                      sequence_1, sequence_2, sequence_3) {
  quantity <- c(quantity_1, quantity_2, quantity_3)
  sequence <- c(sequence_1, sequence_2, sequence_3)

  -sum((quantity - c(0.2, 0.6, 0.8))^2) -
    0.01 * sum((sequence - c(1, 3, 2))^2)
}
```

## Run the complete workflow

```{r complete-workflow, eval=FALSE}
result <- magp_bayes_optimize_from_scratch(
  objective,
  n_initial = 8,
  q = 3,
  model = "2d",
  direction = "maximize",
  n_iter = 5,
  seed = 4,
  design_control = list(
    sequence_maxit = 500,
    quantity_maxit = 500,
    alignment_maxit = 500
  ),
  fit_control = list(
    n_starts = 4,
    workers = 2
  ),
  acquisition_control = list(
    n_starts = 5,
    workers = 2
  )
)

result$best_point
result$best_value
result$history
```

`direction` must state whether smaller or larger responses are preferred. The
returned history keeps the initial and sequential runs together, so each
selection can be reviewed and reproduced.

## Continue from completed experiments

When initial runs and responses already exist, use `magp_bayes_optimize()`
instead. It fits the selected MaGP surrogate, maximizes expected improvement,
evaluates the objective, and refits after each new result.

```{r continue-existing, eval=FALSE}
X_initial <- design$design
y_initial <- apply(X_initial, 1L, function(row) {
  do.call(objective, as.list(row))
})

result <- magp_bayes_optimize(
  objective,
  X = X_initial,
  y = y_initial,
  direction = "maximize",
  n_iter = 5,
  seed = 4
)
```

For laboratory work, do not connect the function directly to equipment until
the proposed point, bounds, units, and safety constraints have been reviewed.
The package selects statistical candidates; it does not replace experimental
oversight.
